The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-03-05 Paper/jaccardbaetica.tex
add new changes

\newcommand{\memo}[2]{\textcolor{#1}{#2}}
\newcommand{\maria}[1]{\memo{red}{#1\\}}
\newcommand{\revise}[1]{\memo{blue}{#1\\}}

\newcommand{\xavi}[1]{\memo{magenta}{XRC: #1\\}}

@@

They could be identified based on three premises: content (olive oil), context (amphora workshop) and subject (individuals involved). On the one hand, it seems that stamps could have been identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}. On the other, they could belong to the owner of the making amphorae workshops or even a group of amphora workers (CITAR). Additionally, they have been related to a quality system \citep{callender} or a production counting system \citep{berni_millet_epigrafianforica_2008}. In any case, the use of these stamps became in a good proxy to define somehow the system of working in the workshops.

@@

The conquest of new provinces allowed to the Roman Empire the arrival of new resources. This lead to a progressive change in the economic and social structure. As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat. The role of the praefectura annonae was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens CITAR (Remesal, 1986; 2011).

@@

The consumption of olive oil in Britannia was residual until the start of the Roman conquest \citep{funari_corpus_1996,
carreras_abastecimiento_2003}.

@@

Most military sites that showed greater similarity in the results were geographically close. We identified two factors in the results: 1) nearby sites tend to share similar stamps and 2) most sites have different stamps. 
By contrast, most sites did not show a strong similarity in stamp correlation . Neither we found a grouping of similar stamps in a specific place. In general, we did not observe any production pattern that indicates a clear organisation between production centres and reception centres. In other words, our results did not indicate the presence of a specific production center from Baetica in Britannia province.  Rather BLA BLA

@@

The similarity of stamps show a results less significant than Britannia province as it can be seen in the dendrogram (Fig. \ref{germap}).

@@

The sites with more similarity in stamps were closely located in limes. In this case, most sites have different stamps  which means that we did not detect any pattern. On the other hand, there seems to be a greater concentration of the seals in areas eminently militarized and close to the German limes. However, It is still no possible to determine with the data that the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.

@@

The results of the case study could be interpreted by several reasons. On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop (CITAR). This hypothesis would explain the fact that we do not find similar stamps in different workshops; however, we found different amphora stamps in the same workshops that they would be barely difficult to assign different owners. Neither a quality distinction does seem to be used to specify the value of the product \citep{callenderindex}. On the other hand, it could be interpreted somehow a batch systematic organisation. Considering that Dressel 20 was not marked in most cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped \citep{berni_millet_epigrafianforica_2008}. This method would be used as an identifier to count the number of amphorae of a branch \citep{juanmorostesis}. This method could also have served to identify different groups of potter workers working in the same amphora workshop. Potters could have marked the amphorae to distinguish different groups working in parallel \citep{li_crossbows_2014}. This it would explain wherefore why we detect different stamps in a same workshop. 
In any case, we do not have enough archaeological evidence that can validate the interpretations presented here and our results are certainly valid only with the context of our case study.

>>>>

\newcommand{\memo}[2]{\textcolor{#1}{#2}}
\newcommand{\maria}[1]{\memo{red}{#1\\}}
\newcommand{\xavi}[1]{\memo{magenta}{XRC: #1\\}}

@@

They could be identified based on three premises: content (olive oil), context (amphora workshop) and subject (individuals involved). On the one hand, it seems that stamps could have been identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}. On the other, they could belong to the owner of the making amphorae workshops or even a group of amphora workers \citep{berni_millet_epigrafianforica_2008}. In any case, the use of these stamps became in a good proxy to define somehow the system of working in the workshops.

@@

The conquest of new provinces allowed to the Roman Empire the arrival of new resources. This lead to a progressive change in the economic and social structure. As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat. The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.

@@

The consumption of olive oil in Britannia was residual until starting the Roman conquest \citep{funari_corpus_1996,
carreras_abastecimiento_2003}.

@@

Most military sites that showed greater similarity in the results were geographically close. We identified two factors in the results: 1) nearby sites tend to share similar stamps and 2) most sites have different stamps. 
By contrast, most sites did not show a strong similarity in stamp correlation. Neither we found a grouping of similar stamps in a specific place. In general, we did not observe any production pattern that indicates a clear organisation between production centres and reception centres. In other words, our results did not indicate the presence of a specific production center from Baetica in Britannia province. Rather it seems that olive oil production was distributed by non specific production centres.

@@

The similarity of stamps shows a results less significant than Britannia province as it can be seen in the dendrogram (Fig. \ref{germap}).

@@

The sites with more similarity in stamps were closely located in limes. In this case, most sites have different stamps  which means that we did not detect any pattern. On the other hand, there seems to be a greater concentration of the seals in areas eminently militarised and close to the German limes. However, It is still no possible to determine with the data that the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.

@@

The results of the case study could be interpreted by several reasons. On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop. This hypothesis would explain the fact that we do not find similar stamps in different workshops; however, we found different amphora stamps in the same workshops that they would be barely difficult to assign different owners. On the other hand, it could be interpreted somehow a batch systematic organisation. Considering that Dressel 20 was not marked in most cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped \citep{berni_millet_epigrafianforica_2008}. This method would be used as an identifier to count the number of amphorae of a branch \citep{juanmorostesis}. This method could also have served to identify different groups of potter workers working in the same amphora workshop. Potters could have marked the amphorae to distinguish different groups working in parallel \citep{li_crossbows_2014}. This it would explain wherefore why we detect different stamps in a same workshop. 
In any case, we do not have enough archaeological evidence that can validate the interpretations presented here and our results are certainly valid only with the context of our case study.
